Neon: Nuclear Norm to Beat Muon

### 2025-05-13 poster/neon_poster.tex
Poster fixed

% This gives good title positioning for a portrait poster.
% Watch out for hyphenation in titles - LaTeX will do it
% but it looks awful.
\begin{textblock}{23}(0,0)
\Title{Neon: Nuclear Norm to Beat Muon}
\end{textblock}

\begin{textblock}{23}(0,0.6)
{
\LARGE
Alexey Kravatskiy, Ivan Kozyrev, Nikolay Kozlov, Alexander Vinogradov, Daniil Merkulov, Ivan Oseledets
}

@@

\begin{textblock}{7.0}(0,1.5)

@@

\begin{textblock}{7.0}(8,1.5)
\hrule\medskip
\Head{Algorithms}\\
Now we can write down pseudocode for both versions of Neon.

@@

\hrule\medskip
\Head{MLP tests}\\
We test Muon, SGD, and Neon on a simple MLP (2 linear layers) that solves CIFAR-10 classification problem. Muon converges faster than Neon and reaches higher accuracy.
\begin{figure}[h]
    \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/loss_vs_epochs_mlp.pdf}}
    \caption{MLP validation loss}
    \label{fig:mlp_loss}
\end{figure}

@@

%\begin{textblock}{7.0}(16,1.5)
%    \begin{figure}[h]
%        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
%        \caption{MLP Accuracy vs wallclock time}
%        \label{fig:mlp_loss}
%    \end{figure}

@@

\begin{textblock}{7.0}(16,1.5)
    \begin{figure}[h]
        \center{\includegraphics[width=0.8\linewidth]{../figures/13may_muon_neon/test_acc_vs_epoch.pdf}}
        \caption{Accuracy vs Epoch for Muon vs Neon}
        \label{fig:mlp_loss}
    \end{figure}

@@

% \begin{figure}[h]
%     \center{\includegraphics[width=0.8\linewidth]{../../figs/mlp24/loss_val.png}}
%     \caption{NanoGPT validation loss}
%     \label{fig:nanogpt_loss}
% \end{figure} 
% \begin{figure}[h]
%     \center{\includegraphics[width=0.8\linewidth]{../../figs/nanogpt/train_loss.pdf}}
%     \caption{NanoGPT train loss}
%     \label{fig:nanogpt_train_loss}
% \end{figure}
\begin{figure}[h]
    \center{\includegraphics[width=0.8\linewidth]{../figs/nanogpt/val_loss.pdf}}
    \caption{NanoGPT validation loss}
    \label{fig:nanogpt_val_loss}
\end{figure}
\medskip
\hrule\medskip
\Head{Conclusion}\\
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\medskip
\hrule\medskip
\Head{Acknowledgements}\\
This material is based upon work supported by the
X Fellowship and my mom.

>>>>

% This gives good title positioning for a portrait poster.
% Watch out for hyphenation in titles - LaTeX will do it
% but it looks awful.
\begin{textblock}{23}(0,-0.25)
\Title{Neon: Nuclear Norm to Beat Muon}
\end{textblock}

\begin{textblock}{23}(0,0.35)
{
\LARGE
Alexey Kravatskiy, Ivan Kozyrev, Nikolay Kozlov, Alexander Vinogradov, Daniil Merkulov, Ivan Oseledets
}

@@

\begin{textblock}{7.0}(0,1.25)

@@

\begin{textblock}{7.0}(8,1.0)
\hrule\medskip
\Head{Algorithms}\\
Now we can write down pseudocode for both versions of Neon.

@@

\hrule\medskip
\Head{MLP tests}\\
We test Muon, SGD, and Neon on a simple MLP (2 linear layers) that solves CIFAR-10 classification problem. Muon converges faster than Neon and reaches higher accuracy.
\begin{figure}[h]
    \center{\includegraphics[width=0.7\linewidth]{../figs/mlp0605/loss_vs_epochs_mlp.pdf}}
    \caption{MLP validation loss}
    \label{fig:mlp_loss}
\end{figure}

\begin{figure}[h]
    \center{\includegraphics[width=0.7\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
    \caption{MLP Accuracy vs wallclock time}
    \label{fig:mlp_loss}
\end{figure}

@@

%\begin{textblock}{7.0}(16,1.5)

\begin{textblock}{7.0}(16,0.75)

@@

    \begin{figure}[h]
        \center{\includegraphics[width=0.75\linewidth]{../figures/13may_muon_neon/test_acc_vs_epoch.pdf}}
        \caption{Accuracy vs Epoch for Muon vs Neon}
        \label{fig:mlp_loss}
    \end{figure}

@@

% \begin{figure}[h]
%     \center{\includegraphics[width=0.8\linewidth]{../../figs/mlp24/loss_val.png}}
%     \caption{NanoGPT validation loss}
%     \label{fig:nanogpt_loss}
% \end{figure} 
% \begin{figure}[h]
%     \center{\includegraphics[width=0.8\linewidth]{../../figs/nanogpt/train_loss.pdf}}
%     \caption{NanoGPT train loss}
%     \label{fig:nanogpt_train_loss}
% \end{figure}
\begin{figure}[h]
    \center{\includegraphics[width=0.8\linewidth]{../figs/nanogpt/val_loss.pdf}}
    \caption{NanoGPT validation loss}
    \label{fig:nanogpt_val_loss}
\end{figure}
\medskip
\hrule\medskip
\Head{Conclusion}\\
A new norm-motivated algorithm proved to be not so effective as Muon, at least from the first sight. However, its updates seem to be much faster to compute, and more so for a large matrices which has not tested yet. We believe that our algorithm is not an alternative to Muon, but probably a more effective tool for solving some other optimization problem. A deep inspection of Nuclear Norm and its properties might give us an insight into Neon's future.

\medskip
%\hrule\medskip
%\Head{Acknowledgements}\\
%We thank Jeremy Bernstein for his idea to use norms to derive a optimizer.
